---
title: "Data Cleaing: Complex Emergencies"
output: github_document
---

The following settings and libraries will be used throughout. 

```{r settings, message = FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(readxl)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# Setting working directory to cleaning 
#setwd('./(1) data_cleaning/')
```

### Pulling data from the API key

```{r message=FALSE, warning=FALSE}
rweb = 
  GET("https://api.reliefweb.int/v1/disasters?appname=rwint-user-0&profile=list&preset=latest&slim=1&query[value]=country.id%3A(120%20OR%20228%20OR%20251%20OR%2044%20OR%2048%20OR%20135%20OR%20147%20OR%20165%20OR%20188%20OR%20212)%20AND%20date.created%3A%5B1980-01-01%20TO%202011-01-01%7D&query[operator]=AND&limit=500") %>% 
   content("text") %>% 
    jsonlite::fromJSON() %>% 
    as_tibble()
```

Api at this [link](https://reliefweb.github.io/search-converter/?appname=rwint-user-0&search-url=https%3A//reliefweb.int/disasters%3Fadvanced-search%3D%2528C120.C228.C251.C44.C48.C135.C147.C165.C188.C212%2529_%2528DA19900101-20101231%2529)

10 Countries ― Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, and Vietnam. 

Country list from ASSOCIATION OF SOUTHEAST ASIAN NATIONS (ASEAN)

Total: 261 Complex Emergencies


### Excel File & Data Cleaning

The variable names in the JSON tibble were difficult so I just copied and pasted it into an excel doc and imported it. 

```{r}
## JSON was difficult to manage, copied the data and imported as an excel file
ce_df = read_xlsx("./datasets/ce_data.xlsx")

## Creating appropriate variables and columns
ce_df = 
ce_df %>% 
  drop_na() %>% 
  
  # Create appropriate date columns
  separate(date, into = c("year_id", "x"), sep = 4) %>% 
  select(-x) %>% 
  
  # Create country column
  separate(name, into = c("location_name", "event"), sep = ":") %>% 
  
  #removing extra month info in description
  separate(event, into = c("event", "r"), sep = "-") %>% 
  separate(location_name, into = c("location_name", "s"), sep = "-" ) %>% 
  select(-r, -s) %>% 
  
  # Renaming various Lao country names to Lao PDR
  mutate(location_name = replace(location_name, str_detect(location_name, "Lao"), "Lao PDR")) %>%
  
  # Filtering only by events located within a specific country (to remove regional typhoons)
  filter(location_name %in% c("Cambodia", "Indonesia", "Lao PDR", 
                              "Malaysia", "Myanmar", "Philippines", "Singapore", 
                              "Thailand", "Viet Nam", "Southeast Asia", "South Asia")) %>% 
  
  # Creating a categorical level variable for total number of CEs
  mutate(
    cat_ce = case_when(
      location_name %in% c("Brunei Darussalam", "Viet Nam", "Singapore") ~ "None",
      location_name %in% c("Lao PDR", "Malaysia", "Cambodia") ~ "Low",
      location_name %in% c("Thailand", "Myanmar") ~ "Medium",
      location_name %in% c("Philippines", "Indonesia") ~ "High")
    )


#Export
write_csv(ce_df, "./datasets/clean_ce_data.csv")


# Creating a dataset with summed values to determine levels
ce_sum = 
  ce_df %>% 
    group_by(location_name) %>% 
    summarize(
      total_ce = n())


  
## Note: No CEs recorded in Brunei, Vietnam, Singapore
## Categories:
##    None:Brunei, Vietnam, Singapore [0]
##    Low: Lao PDR, Malaysia, Cambodia [5-10]
##    Medium: Thailand, Myanmar [16 - 23]
##    High: Philippines, Indonesia [92 - 108]
```





