---
title: "Data Cleaing: IHME Fertility estimates"
output: github_document
---

```{r settings, message = FALSE}
library(tidyverse)
library(dplyr)
```

### Reading datasets 
The following metrics are considered for this project

* Total Fertility Rate (TFR)
* Net Reproductive Rate (NRR)
* Crude Birth Rate (CBR)

ASEAN countries considered here are

* Brunei Darussalam
* Cambodia 
* Indonesia 
* Lao People's Democratic Republic 
* Malaysia
* Myanmar 
* Philippines 
* Singapore
* Thailand
* Viet Nam

```{r, message = FALSE}
asean_countries = 
  c("Brunei Darussalam", "Cambodia", "Indonesia",  "Lao People's Democratic Republic", "Malaysia", "Myanmar",  "Philippines",  "Singapore", "Thailand", "Viet Nam")

# Total Fertility Rate
tfr_df =
  read_csv('./datasets/GBD_datasets/IHME_GBD_2019_FERTILITY_1950_2019_TFR_Y2020M08D05.CSV') %>%
  filter(location_name %in% asean_countries | location_name == 'Global',
         year_id >= 1980 & year_id <= 2010) %>%
  select(-c(sex, metric_id, measure_id, metric_name, age_group_id, sex_id, location_id)) %>%
  distinct() %>%
  rowwise() %>%
  mutate(measure_name = 'TFR')

# Net Reproductive Rate
nrr_df =
  read_csv('./datasets/GBD_datasets/IHME_GBD_2019_FERTILITY_1950_2019_NRR_Y2020M08D05.CSV') %>%
  filter(location_name %in% asean_countries | location_name == 'Global',
         year_id >= 1980 & year_id <= 2010) %>%
  select(-c(sex, metric_id, measure_id, metric_name, age_group_id, sex_id, location_id)) %>%
  rowwise() %>%
  mutate(measure_name = 'NRR')

# Crude Birth Rate
cbr_df = 
  read_csv('./datasets/GBD_datasets/IHME_GBD_2019_FERTILITY_1950_2019_CBR_Y2020M08D05.CSV') %>%
  filter(location_name %in% asean_countries | location_name == 'Global',
         year_id >= 1980 & year_id <= 2010) %>%
  select(-c(sex, metric_id, measure_id, metric_name, age_group_id, sex_id, location_id)) %>%
  rowwise() %>%
  mutate(measure_name = 'CBR')

```

Each dataset includes the following 7 columns

* location_name
* age_group_name
* year_id
* measure_name
* val
* upper
* lower 

### Combining all dataframes together

```{r}
gbd_fertility_df = rbind(tfr_df, nrr_df, cbr_df) %>%
  write_csv('./datasets/gbd_fertility.csv')
```

The merged dataset contains `r nrow(gbd_fertility_df)` rows and `r ncol(gbd_fertility_df)` columns in total. It includes 10 Southeast Asian countries along with global estimate for each metric calculated between 1980 and 2010.

